---
title: "p8105_hw2_wx2233"
author: "Weijia Xiong"
date: "9/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1
### Read and clean the Mr. Trash Wheel sheet
```{r}
Wheel_data = 
 readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet = 1) %>%  #import data from sheet 1
  janitor::clean_names() %>%  #use reasonable variable names
  select(everything(),waste_energy_plant = x15) %>% 
  drop_na(dumpster) %>% #omit rows that do not include dumpster-specific data
  mutate(
    sports_balls = as.integer(round(sports_balls))
  )

Wheel_data
```

### Read and clean precipitation data for 2017 and 2018

```{r}
pre_data_2017 = 
 readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet = "2017 Precipitation",range = "A3:B14",col_names = c("month","precipitation")) %>%#import data from sheet 2017 Precipitation 
  drop_na(precipitation) %>% 
  mutate(
    year = rep("2017")
  )
  

pre_data_2018 = 
 readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet = "2018 Precipitation",range = "A3:B14",col_names = c("month","precipitation")) %>%#import data from sheet 2018 Precipitation 
  drop_na(precipitation) %>% 
  mutate(
    year = rep("2018")
  )

combine_pre_data = 
  bind_rows(pre_data_2017,pre_data_2018) %>% 
  mutate(
    month = month.abb[as.numeric(month)],
    precipitation = as.numeric(precipitation)
  )

combine_pre_data 


combine_pre_tidy_data = 
  combine_pre_data %>% 
  pivot_wider(names_from = year,
              values_from = precipitation)
sum_pre_2018 = sum(pull(combine_pre_tidy_data,"2018"),na.rm = TRUE)
median_sb_2017 = 
  filter(Wheel_data,year == "2017") %>% 
  pull(sports_balls) %>% 
  median()
```

Wheel data has 285 observations and 15 variables:

* dumpster: the id of dumpster
* month: the month of observation
* year: the year of observation
* date: the date of observation
* weight_tons: the whole weight of waste
* volume_cubic_yards: the volume of cubic yards waste
* plastic_bottles, polystyrene, chips_bag, glass_bottles, grocery_bags, chip_bags, sports_balls, homes_powered are different types of waste.

Combination of precipitation data has 3 variables and 13 observations.
The second and the third column represent the precipitation for 2017 and 2018.

* The total precipitation in 2018 is `r sum_pre_2018`
* The median number of sports balls in a dumpster in 2017 is `r median_sb_2017`


## Problem 2

### pols-month data
```{r}
#pols_month
pols_month = read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    month = month.abb[as.numeric(month)],
    prez_gop = as.character(prez_gop),
    president = recode(prez_gop,  "0" = "dem",  "1" = "gop")  # create a president variable
  ) %>% 
  select(everything(),ends_with("gop"), ends_with("dem"), -prez_dem, -prez_gop, -day)

pols_month
```

### snp data
```{r}
#snp_data
snp_data = read_csv("./data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day","year"), sep = "/") %>% 
  arrange(year, month) %>% 
  mutate(
    month = month.abb[as.numeric(month)]
  ) %>% 
  select(year, month, everything(),-day)

snp_data
```

## unemployment data
```{r}
#unemployment
unemployment_data = read_csv("./data/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "unemployment") %>% 
  rename(year = Year) %>% 
  mutate(
    year = as.character(year)
  ) %>% 
  drop_na(unemployment)

unemployment_data

```



### Merge and join
```{r}
join_data = left_join(pols_month, snp_data, by = c('year','month')) # can join by multiple variables 

final_join = left_join(join_data, unemployment_data,by = c('year','month'))

final_join
```

The "pols_month" contains 822 observations (from Jan, 1947 to Jun, 2015) of 9 variables: 

* year: the year of observation:
* month: the month of observation
* gov_gop: the number of republican governors on the associated month of the associated year
* sen_gop: the number of republican senators on the associated month of the associated year
* rep_gop: the number of republican representatives on the associated month of the associated year
* gov_dem: the number of democratic governors on the associated month of the associated year
* sen_dem: the number of democratic senators on the associated month of the associated year
* rep_dem: the number of democratic representatives on the associated month of the associated year
* president: indicator of whether the president was republican or democratic the associated month of the associated year (1 = republican, 0 = democratic)


The "snp_data" contains 787 observations (from Jan, 1950 to Jun, 2015) of 3 variables: 

* year: the year of observation
* month: the month of observation
* close: the closing values of the S&P stock index on the associated month of the associated year


The "unemployment_data" contains 816 observations (from Jan, 1948 to Jun, 2015) of 3 variables: 

* year: the year of observation
* month: the month of observation
* unemployment: percentage of unemployment in the associated month of the associated year



## Problem 3

```{r}

Popular_baby = read_csv("./data/Popular_Baby_Names.csv")  %>% 
  janitor::clean_names() %>% 
  select(year_of_birth, childs_first_name,everything()) %>% 
  mutate(
     gender = str_to_lower(gender),
     ethnicity = str_to_lower(ethnicity)
  ) %>% 
  distinct()
Popular_baby
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. Produce a similar table showing the most popular name among male children over time.

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis)

### Compare the rank of Olivia
```{r}
Olivia = Popular_baby %>% 
  filter(childs_first_name == "Olivia", gender == "female") %>% 
  select(-count) %>% 
  pivot_wider(
    names_from = ethnicity,
    values_from = rank
) %>% 
  arrange(year_of_birth,desc(year_of_birth))
Olivia
```

### Scatter plot for white non hispanic male in 2016
```{r}
male_name_2016_wnh = Popular_baby %>% 
  filter(gender == "male", year_of_birth == "2016", ethnicity == "white non hispanic")

male_name_2016_wnh %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point() +
  geom_smooth(se = FALSE)

```


